Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-29 introduction/Subsections/Introduction.tex
Updates from Overleaf

MIR is a research area that has emerged in the last decades as an interdisciplinary field combining engineering, musicology, and neuroscience. MIR aims to develop algorithms and techniques to automatically analyze, classify, and retrieve musical information from mainly audio signals. I believe that traditionally, MIR research has primarily concentrated on technical aspects, such as Digital Signal Processing (DSP), or pattern recognition. As a result, other relevant domains, including perception, cognition, music theory, aural skills, and musicology, along with its subfields, such as the symbolic domain, orchestration, and automatic transcription, have often been overlooked. However, it is worth noting that there has been a significant shift in this tendency recently.

The reason behind this is non-other but the implementation complexity for such heuristics and abstract fields of study.

>>>>

MIR is a research area that has emerged in the last decades as an interdisciplinary field combining engineering, musicology, and neuroscience. MIR aims to develop algorithms and techniques to automatically analyze, classify, and retrieve musical information from mainly audio signals. I believe that traditionally, MIR research has primarily concentrated on technical aspects, such as Digital Signal Processing (DSP) or pattern recognition. As a result, other relevant domains, including perception, cognition, music theory, aural skills, and musicology, along with its subfields, might have been overlooked.

It is worth mentioning that there has been a significant shift in this tendency on retrieval tasks in the symbolic and musicological domain, such as powerful open-source analysis frameworks \cite{ScottCuthbertChristopherArizaMusic21:Data}, MIDI to proper sheet music syntax \cite{Suzuki2021ScoreRepresentation}, projective orchestration \cite{Crestel2018AOrchestration}, or automatic lead sheet generation \cite{WeilAutomaticSignals} to name a few.

The reason behind this is non-other but the implementation complexity for such heuristics and abstract fields of study. Closing the gap between human perception, environmental and cultural factors, and computational solutions to mimic the formers is an already tackled topic \cite{Large1994ResonanceMeter}\cite{Mullensiefen2003MeasuringJudgments}\cite{Sears2014PerceivingCadence}\cite{Thaut2014HumanPatter}, yet remains a vast open challenge.

### 2023-03-29 introduction/Subsections/Motivation.tex
Updates from Overleaf

\section{Motivation}

@@

Let's pick a core MIR task as a sample to display such complexity: Music Structure Analysis (MSA), an interdisciplinary field that aims to understand the structure of music\cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

@@

As a music lover, professional musician, and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them using DN.  I am eager to learn how DNNs listen to music. I am deeply interested in learning how to retrieve embedded information within music.

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees, and the tonal function.}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

@@

>>>>

\section{Motivation}
As a music lover, professional musician, and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them. I am deeply interested in learning how to retrieve useful embedded information within the music.

\subsection{Beyond Digital Signal Processing (DSP)}

@@

Let's pick a core MIR task as a sample to display such complexity: Music Structure Analysis (MSA), an interdisciplinary field that aims to understand the structure of music\cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

@@

@@

\begin{figure}[h]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees, and the tonal function.}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

@@

Schenkerian analysis is a method of analyzing tonal music, focusing on hierarchical relationships between musical elements. The theory is based on the ideas of Heinrich Schenker, who sought to show that free composition was an elaboration of strict composition, particularly two-voice counterpoint. Schenker's theory includes several key concepts:

Prolongational levels: Hierarchically organized levels of elaboration, with each level representing a new freedom with respect to the rules of strict composition.
Harmony: The basic component is the Stufe (scale degree), and Schenker's theory is monotonal, meaning that the work as a whole projects a single key.
Counterpoint and voice-leading: Schenker emphasizes the importance of two-voice counterpoint and melodic fluency (stepwise motion) in both strict and free composition.
Ursatz (fundamental structure): The underlying structure from which a work originates, consisting of a fundamental line (Urlinie) and bass arpeggiation (Bassbrechung).
Schenkerian analysis is about showing how each work elaborates the background structure in a unique way, determining its identity and meaning. While the theory has been criticized for reducing all tonal works to a few background structures, Schenker argues that the focus is on individual elaboration.

### 2023-03-29 introduction/Subsections/Objectives.tex
Updates from Overleaf

This work aims to uncover a generalization of the fundamental tonal-working structure of western-tradition music by learning and embedding space used to facilitate its application in downstream MIR tasks.

Understanding how music is organized and composed is critical, and music theory plays a vital role.

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the level of musical abstraction and the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

>>>>

This work aims to uncover a generalization of the fundamental tonal-working structure of western-tradition music by learning and embedding space used to facilitate its application in downstream MIR tasks. Understanding how music is organized and composed is critical, and music theory plays a vital role.

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages \cite{}. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the level of musical abstraction and the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.
